Online Balanced Graph Partitioning: Optimizations for Large Clusters

### 2020-05-07 paper/disc20clustering.tex
restructuring the proof
comments

 \newcommand\mahmoud[1]{\color{orange}\textbf{Mahmoud: #1}\\\color{black}}
 \newcommand\stefan[1]{\color{blue}\textbf{\\ Stefan: #1}\color{black}}
 \newcommand\maciek[1]{\color{brown}\textbf{(Maciek: #1)}\color{black}}
 %\newcommand\maciek[1]{}

@@

\begin{proof}
	Fix any online algorithm \ALG{}.
	Let $I(C)$ denote the cluster where nodes of ground set $C$ are located initially.
	If $C$ contains nodes that originate from different clusters, $I(C)$ is undefined.
	Throughout the construction, we maintain ground sets for each node.
	Initially all nodes are \emph{alone} (they belong to their own trivial ground sets).

	First, we create a ground set of size $k-1$ on an arbitrarily chosen cluster $S$.
	%We never dismiss any constructed ground set.
	Each cluster hosts exactly $k$ nodes, and in any configuration a single alone node resides in $S$.

	Then, we add the alone node from $S$ to a larger ground set to force its eviction.
	Precisely,
	we create a ground set $\{x_0, y_0\}$, 
	where $x_0$ is the alone node from $S$ and $y_0$ is an arbitrary node that does not reside in $S$.
	\ALG has them split, and the adversary produces external requests until they are collocated.
	The cluster $S$ already contains a ground set of size $k-1$, hence \ALG must collocate $\{x_0, y_0\}$ in another cluster.
	To preserve a feasible partition of nodes into clusters after collocating $\{x_0, y_0\}$,
	\ALG must replace $x_0$ with another single node.

	We continue adding the node that is currently alone in $S$ to a larger ground set.
	Precisely, for an alone node $x$ from~$S$, we merge the ground sets of $x$ and the largest ground set $C$ s.t. $C \neq \{x_0,y_0\}$ and $I(C) = I(x)$.
	This way, $x$ is evicted similarly to $x_0$, and we continue creating such ground sets.
	We describe the termination condition shortly.
	(Note that these ground sets contain nodes that originate from the same cluster, whereas the nodes $x_0$ and $y_0$ originate from different clusters.)

	We claim that expanding ground sets in this way can continue as long as at least $\ell$ alone node exists.
	Fix the moment just before the creation of a ground set with the alone node $z$ that currently resides in $S$.
	We show that if at least $\ell$ alone nodes exists (including $z$), then there exists a perfect partition of ground sets after expanding a ground set with $z$.

	To show this claim, we consider a configuration, where the largest ground set of each cluster originates from it (excluding the ground set $\{x_0, y_0\}$).
	In this configuration, each cluster contains:  at most one ground set larger than $1$, possibly some alone nodes, and possibly the ground set $\{x_0, y_0\}$.
	If $I(z)$ contains an alone node, then we simply obtain a~perfect partition by swapping it with $z$.
	In the following we assume that no alone nodes exist on $I(z)$.
	The largest ground set of $I(z)$ is at most $k-1$, because $z$ is on $S$.
	Therefore $\{x_0, y_0\}$ is present on $I(z)$, and we must evict it to make space for $z$.
	To find the cluster for $\{x_0, y_0\}$, observe that we have at most $\ell-1$ clusters with alone nodes, and we have at least $\ell$ alone nodes --- hence a cluster with two alone nodes exists.
	To find a feasible partition, we swap $\{x_0, y_0\}$ with these two collocated alone nodes, and then we swap an alone node from $I(z)$ with $z$.

@@

Observe that for requests in
$R_t := \sigma_{\{u,v'\}}(t'',t'] \cup \sigma_{\{u,v\}}(t',t]$ 
and the entire duration of the input sequence (i.e., $(0,m]$),
OPT must pay a total cost
%We denote the cost w.r.t~$r_t$ by
$\mathit{Cost}_{\mathit{OPT}} (R_t) \geq 
	c_{R_t} := \min{ \{ 2\lambda, \lambda+\alpha, 2\alpha \} }$,
for serving (some of) these requests remotely or (and) collocating (one of) the pairs,
regardless of its choices.
Then the cost per pair, in particular the cost w.r.t.~$\{u,v\}$, is
$\mathit{Cost}_{\mathit{OPT}} (\sigma_{\{u,v\}}(t',t]) \geq c_{R_t} / 2$.
%
However,
we can charge this cost to OPT only at $t=t_0$,
i.e.~the first time Alg collocates $\{u,v\}$,
and not at any consequent time $\tilde{t} > t_0$ that Alg collocates $\{u,v\}$.
%	i.e.,
%	when Alg collocates the two pairs a second time.
That is,
we charge OPT a cost incurred by requests in
$\sigma_{\{u,v'\}}(t'',t']$ or $\sigma_{\{u,v\}}(t',t]$,
if and only if  OPT incurs the cost during  intervals $(t'', t']$ or $(t', t]$,
respectively.	
Latter,
this ensures us that we do not charge OPT for a cost more than twice.
%%
Based on this observation,
in the proof that follows,
we distinguish three cases of OPT's choices
while it serves requests in	$R_t $.

@@

	Next, we sum up all costs incurred by OPT.
	Observe that by definition,
	$R_{t'} \cap R_t = \sigma_{\{u,v'\}}[t'',t']$.
	This means,
	the cost of r.h.s.~to OPT is counted at most twice in each of  the expressions
	$\mathit{Cost}_{\mathit{OPT}} (R_{t'}) + \mathit{Cost}_{\mathit{OPT}} (R_t)$
	and  
	$\mathit{Cost}_{\mathit{OPT}} (R_{t'}) + \mathit{Cost}_{\mathit{OPT}} (R_{\{u,v\}})$.
	Hence,
	for all collocations performed by Alg,
	and for final costs at $t=m$,
	OPT pays at least 
	$\frac{1}{2}(
	\sum_{ t \in [1,m] } \mathit{Cost}_{\mathit{OPT}} (R_t) +
	\sum_{\{u,v\}} \mathit{Cost}_{\mathit{OPT}} (R_{\{u,v\}})
	) $.
	Then,
	the total cost to OPT is

>>>>

 \newcommand\mahmoud[1]{\color{orange}\textbf{Mahmoud: #1~}\color{black}}
 \newcommand\stefan[1]{\color{blue}\textbf{\\ Stefan: #1}\color{black}}
 \newcommand\maciek[1]{\color{brown}\textbf{(Maciek: #1)}\color{black}}
 %\newcommand\maciek[1]{}

@@

\begin{proof}
	Fix any online algorithm \ALG{}.
	Let $I(C)$ denote the cluster where nodes of the ground set $C$ are located initially.
	If $C$ contains nodes that originate from different clusters, $I(C)$ is undefined.
	\mahmoud{i.e., $I(C):=...$formalized with notations}.
	Throughout the construction, we maintain ground sets for each node.
	Initially all nodes are \emph{alone} (they belong to their own trivial ground sets).
\mahmoud{isolated would be a better word for "alone"}
	First, we create a ground set of size $k-1$ on an arbitrarily chosen cluster $S$.
	\mahmoud{$\mathcal{S}$ is better for a set of sets.}
	%We never dismiss any constructed ground set.
	Each cluster hosts exactly $k$ nodes, and in any configuration a single alone node resides in $S$. \mahmoud{an isolated node is collocated with  this GS in $\mathcal{S}$. We denote the first such node by $x_0$.}

	Then, we add the alone node from $S$ to a larger ground set to force its eviction.
	Precisely,
	we create a ground set $\{x_0, y_0\}$, 
	where $x_0$ is the alone node from $S$ and $y_0$ is an arbitrary node that does not reside in $S$.
	\ALG has them split, and the adversary issues external requests until \ALG collocates them.
	The cluster $S$ already contains a ground set of size $k-1$, hence \ALG must collocate $\{x_0, y_0\}$ in another cluster.
	To preserve a feasible partition of nodes into clusters after collocating $\{x_0, y_0\}$,
	\ALG must replace $x_0$ with another single node.
	\mahmoud{In any new partition, \ALG must replace $x_0$..., otherwise it is not a feasible partition.}

	We continue adding the node that is currently alone in $S$ to a larger ground set.
	Precisely, for an alone node $x$ from~$S$, we merge the ground sets of $x$ and the largest ground set $C$ s.t. $C \neq \{x_0,y_0\}$ and $I(C) = I(x)$.
	\mahmoud{$x_i$?}
	This way, $x$ is evicted similarly to $x_0$, and we continue creating such ground sets.
	We describe the termination condition shortly.
	(Note that these ground sets contain nodes that originate from the same cluster, whereas the nodes $x_0$ and $y_0$ originate from distinct clusters.)

	We claim that expanding ground sets in this way can continue as long as at least $\ell$ alone node exists.
	Fix the moment just before the creation of a ground set with the alone node \mahmoud{$x_{t^*}$?} that currently resides in $S$.
	\mahmoud{ALG might change $S$. We can say, "currently collocated with the...".}
	We show that if at least $\ell$ alone nodes exists (including $z$), then there exists a perfect partition of ground sets after expanding a ground set with $z$.
	\mahmoud{the termination condition is needed before the claim statement.}
	To show this claim, we consider a configuration, where the largest ground set of each cluster originates from it (excluding the ground set $\{x_0, y_0\}$).
	In this configuration, each cluster contains:  at most one ground set larger than $1$, possibly some alone nodes, and possibly the ground set $\{x_0, y_0\}$.
	If $I(z)$ contains an alone node, then we simply obtain a~perfect partition by swapping it with $z$.
	In the following we assume that no alone nodes exist on $I(z)$.
	The largest ground set of $I(z)$ is at most $k-1$, because $z$ is on $S$.
	Therefore $\{x_0, y_0\}$ is present on $I(z)$, and we must evict it to make space for $z$.
	To find the cluster for $\{x_0, y_0\}$, observe that we have at most $\ell-1$ clusters with alone nodes, and we have at least $\ell$ alone nodes --- hence a cluster with two alone nodes exists.
	To find a feasible partition, we swap $\{x_0, y_0\}$ with these two collocated alone nodes, and then we swap an alone node from $I(z)$ with $z$.

@@

Observe that for requests in
$R_t := \sigma_{\{u,v'\}}(t'',t'] \cup \sigma_{\{u,v\}}(t',t]$ 
and the entire duration of the input sequence (i.e., $(0,m]$),
OPT must pay a total cost
%We denote the cost w.r.t~$r_t$ by
$\mathit{Cost}_{\mathit{OPT}} (R_t) \geq 
	c_{R_t} := \min{ \{ 2\lambda, \lambda+\alpha, 2\alpha \} }$,
for serving (some of) these requests remotely or (and) collocating (one of) the pairs,
regardless of its choices.
Then the cost per pair, in particular the cost w.r.t.~$\{u,v\}$, is
$\mathit{Cost}_{\mathit{OPT}} (\sigma_{\{u,v\}}(t',t]) \geq c_{R_t} / 2$.
%
However,
we can charge this cost to OPT only at $t=t_0$,
i.e.~the first time Alg collocates $\{u,v\}$,
and not at any consequent time $\tilde{t} > t_0$ that Alg collocates $\{u,v\}$.
%	i.e.,
%	when Alg collocates the two pairs a second time.
That is,
we charge OPT a cost incurred by requests in
$\sigma_{\{u,v'\}}(t'',t']$ or $\sigma_{\{u,v\}}(t',t]$,
if and only if  OPT incurs the cost during  intervals $(t'', t']$ or $(t', t]$,
respectively.	
Latter,
this ensures us that we do not charge OPT for a cost more than twice.
%%
We use notations defined so far in the proof that follows.

@@

	Next, we sum up all costs incurred by OPT.
	By definitions of $R_t$ and $R_{\{u,v\}}$, we have either
	$R_{t'} \cap R_t = \sigma_{\{u,v'\}}[t'',t']$ or
	$R_{t'} \cap R_{\{u,v\}} = \sigma_{\{u,v'\}}[t'',t']$
	(\mahmoud{See figure \ref{fig:doubleCount}}). 
	This means,
	$\mathit{Cost}_{OPT} ( \sigma_{\{u,v'\}}[t'',t'])$
	is counted at most twice in each of  the expressions
	$\mathit{Cost}_{\mathit{OPT}} (R_{t'}) + \mathit{Cost}_{\mathit{OPT}} (R_t)$
	and  
	$\mathit{Cost}_{\mathit{OPT}} (R_{t'}) + \mathit{Cost}_{\mathit{OPT}} (R_{\{u,v\}})$.
	Hence,
	for all collocations performed by Alg,
	and for final costs at $t=m$,
	OPT pays at least 
	$\frac{1}{2}(
	\sum_{ t \in [1,m] } \mathit{Cost}_{\mathit{OPT}} (R_t) +
	\sum_{\{u,v\}} \mathit{Cost}_{\mathit{OPT}} (R_{\{u,v\}})
	) $.
	Then,
	the total cost to OPT is
